---
title: "04_celltyping"
format: html
editor: visual
---

## Cell type assignment

```{r}
rm(list = ls())

```

#### Librairy

```{r, massage = False}

library(dplyr)
library(Seurat)
library(patchwork)
library(here)
library(SingleR)
library(celldex)
library(SeuratData)
library(dplyr)
library(tidyr)
library(multtest)
library(metap)
library(tibble)
library(purrr)
library(copykat)
```

#### Data Load

```{r, massage = False}

snRNAseq <- readRDS("../data/04_RNA_seurat")

```

Dim plot

```{r}

DimPlot(snRNAseq, reduction = 'umap', group.by = 'RNA_snn_res.0.1')

```

#### Marker genes cell typing

find conserved markger genes acrocces clusters.

```{r}
Idents(snRNAseq) <- "RNA_snn_res.0.1"
DefaultAssay(snRNAseq) <- "RNA"
annotations <- read.csv("/home/projects/22102_single_cell/day3/annotation.csv")
```

```{r}


cluster0_conserved_markers <- FindConservedMarkers(snRNAseq,
                              ident.1 = 0,
                              grouping.var = "condition",
                              only.pos = TRUE,min.pct = 0.25,  min.diff.pct = 0.25,
                      logfc.threshold = 0.25)

#Load the table with gene descriptions to help you identify to what cell types the genes in the cluster can correspond to



# Combine markers with gene descriptions 
cluster0_ann_markers <- cluster0_conserved_markers  |>  
                rownames_to_column(var="gene") |> 
                left_join(y = 
                            unique(
                              annotations[,c("gene_name", "description")]),
                          by = c("gene" = "gene_name"))

cluster0_ann_markers$gene
```

```{r}

snRNAseq <- RenameIdents(object = snRNAseq, 
                                "0" = "epithelial cells",
                                "1" = "CD8+ T cells",
                                "2" = "CD4+ T cells",
                                "3" = "NK cells",
                                "4" = "Monocyte",
                                "5" = "Monocyte",
                                "6" = "T reg",
                                "7" = "DC",
                                "8" = "B cells",
                                "9" = "Unknown",
                                "10" = "Platelet",
                                "11" = "Hematopoietic stem cells",
                                "12" = "Plasma Cells")

# Now, you can view the updated metadata
head(snRNAseq@meta.data)
```

```{r}
# visualize data

snRNAseq$cell_type_1 <- Idents(snRNAseq)

DimPlot(snRNAseq, reduction = 'umap', label = TRUE)
```

#### Singler R cell typing

```{r}

ref <- HumanPrimaryCellAtlasData()
```

```{r}
serat_ref <- readRDS("/home/projects/22102_single_cell/day3/pbmcsca.rds")

```

```{r}
ref <- as.SingleCellExperiment(serat_ref, assay = "RNA")

pbmc_counts <- GetAssayData(snRNAseq, assay = "RNA", layer = 'counts')

pred <- SingleR(test = pbmc_counts,
                ref = ref,
                labels = ref$CellType,
                de.method="wilcox")

snRNAseq$cell_type_2 <- pred$labels[match(rownames(snRNAseq@meta.data), rownames(pred))]
```

#### 

#### CopyKat

Extrac raw counts from the data

```{r, massage = False}

exp.rawdata <- as.matrix(snRNAseq@assays$RNA$counts)

```

Counts chormos by genes. Some parameters are\
ngene.chr: is the mini number of genes to represent chomoromes, lower number will include more cells\
win.size: number of genes pr seqment\
Ks: seqment cuts senesityvity\
n.cores: numbers of cores to run in paralize\
low.DR: minum numbers of genes in fraction for something with smoothing\
upper.DR: minumun number of genes in fraction for someting with seqments

```{r}
copykat.test <- copykat(rawmat=exp.rawdata, id.type="S", ngene.chr=5, win.size=25, KS.cut=0.1, sam.name="test", distance="euclidean", norm.cell.names="",output.seg="FLASE", plot.genes="TRUE", genome="hg20",n.cores=1)

```

```{r}
pred.test <- data.frame(copykat.test$prediction)
pred.test <- pred.test[-which(pred.test$copykat.pred=="not.defined"),]  ##remove undefined cells
CNA.test <- data.frame(copykat.test$CNAmat)
```

```{r}
my_palette <- colorRampPalette(rev(RColorBrewer::brewer.pal(n = 3, name = "RdBu")))(n = 999)

  chr <- as.numeric(CNA.test$chrom) %% 2+1
  rbPal1 <- colorRampPalette(c('black','grey'))
  CHR <- rbPal1(2)[as.numeric(chr)]
  chr1 <- cbind(CHR,CHR)

  rbPal5 <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "Dark2")[2:1])
  com.preN <- pred.test$copykat.pred
  pred <- rbPal5(2)[as.numeric(factor(com.preN))]

  cells <- rbind(pred,pred)
  col_breaks = c(seq(-1,-0.4,length=50),seq(-0.4,-0.2,length=150),seq(-0.2,0.2,length=600),seq(0.2,0.4,length=150),seq(0.4, 1,length=50))

heatmap.3(t(CNA.test[,4:ncol(CNA.test)]),dendrogram="r", distfun = function(x) parallelDist::parDist(x,threads =4, method = "euclidean"), hclustfun = function(x) hclust(x, method="ward.D2"),
            ColSideColors=chr1,RowSideColors=cells,Colv=NA, Rowv=TRUE,
            notecol="black",col=my_palette,breaks=col_breaks, key=TRUE,
            keysize=1, density.info="none", trace="none",
            cexRow=0.1,cexCol=0.1,cex.main=1,cex.lab=0.1,
            symm=F,symkey=F,symbreaks=T,cex=1, cex.main=4, margins=c(10,10))

  legend("topright", paste("pred.",names(table(com.preN)),sep=""), pch=15,col=RColorBrewer::brewer.pal(n = 8, name = "Dark2")[2:1], cex=0.6, bty="n")
```

```{r}
tumor.cells <- pred.test$cell.names[which(pred.test$copykat.pred=="aneuploid")]
tumor.mat <- CNA.test[, which(colnames(CNA.test) %in% tumor.cells)]
hcc <- hclust(parallelDist::parDist(t(tumor.mat),threads =4, method = "euclidean"), method = "ward.D2")
hc.umap <- cutree(hcc,2)

rbPal6 <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "Dark2")[3:4])
subpop <- rbPal6(2)[as.numeric(factor(hc.umap))]
cells <- rbind(subpop,subpop)

heatmap.3(t(tumor.mat),dendrogram="r", distfun = function(x) parallelDist::parDist(x,threads =4, method = "euclidean"), hclustfun = function(x) hclust(x, method="ward.D2"),
            ColSideColors=chr1,RowSideColors=cells,Colv=NA, Rowv=TRUE,
            notecol="black",col=my_palette,breaks=col_breaks, key=TRUE,
            keysize=1, density.info="none", trace="none",
            cexRow=0.1,cexCol=0.1,cex.main=1,cex.lab=0.1,
            symm=F,symkey=F,symbreaks=T,cex=1, cex.main=4, margins=c(10,10))

  legend("topright", c("c1","c2"), pch=15,col=RColorBrewer::brewer.pal(n = 8, name = "Dark2")[3:4], cex=0.9, bty='n')

```

```{r}

saveRDS(snRNAseq, file = "../data/05_RNA_seurat")
```

#### 
